|

### 2021-06-01 article/article.Rmd
final corrections

Aggregate results indicate that legislature size has no significant impact on
public spending. All of our main meta-analysis estimates show that the overall
effect is not statistically different from zero, thus confirming the
conflicting findings reported by the literature. However, when we look only at
articles that employ regression discontinuity designs (RDDs), we see that all
four papers included in our sample suggest that a higher number of legislators
leads to lower public spending [@debenedetto2018effect; @hohmann2017effect;
@lewis2019legislature; @petterssonlidbom2012size]. In this regard, it is
possible that methodological choices partially explain the divergent results
observed in the literature, as articles that use RDDs consistently point to the
same direction. One limitation of this finding is that these studies only test
the impact of lower house size on the natural logarithm of expenditure per
capita, thus it remains unclear whether the results hold with other variables.

The meta-regressions provide additional evidence that our study sample is
highly heterogeneous and that effect sizes differ substantially according to
study specifications. When using an extended sample of `r dim(fulldat)[1]`
coefficients, we find that unicameralism is associated with higher public
spending, as predicted by the "law of $1/n$". Since most unicameral
legislatures in our sample are local governments (municipalities or districts),
this result supports the predictions of @weingast1981political. Moreover,
our meta-regressions indicate that larger upper chambers spend more in terms of
per capita expenditure than lower chambers, a result that also appears in the
binomial tests. Non-majoritarian voting systems decrease government spending as
a percentage of GDP, but they do not have a relevant impact on the other
measures of public expenditure, following the idea that the $1/n$ effect grows
weaker as the empirical cases distance from the original definition of the law.
Finally, meta-regression results confirm that regression discontinuity designs
reduce public spending estimates.

@@

In the second round, we did not require articles to cite @weingast1981political
and used a keyword-based query on Google Scholar to broaden the scope of the
first search. The search string contained terms strongly associated with the
literature on the "law of $1/n$" and was as follows: `("upper chamber size" OR
"lower chamber size" OR "council size" OR "parliament size" OR "legislature
size" OR "number of legislators" OR "legislative size") AND ("spending" OR
"expenditure" OR "government size")`. We again restricted the search to
articles written in English which employed quantitative methods. This search
added two new results to our sample [@coate2011government;
@debenedetto2018effect], but neither of them included variables beyond the six
measures we had previously identified. In the third search round, we looked
into the personal webpages of every author whom we had already included in our
sample. The purpose of this manual search was to assess whether there was any
working paper or unpublished manuscript that we had missed in the two former
queries. We did not find any new article that satisfied the inclusion criteria
in this search. The full list of excluded records is available for online
consultation in the replication materials. Combined, the three searches
produced a dataset of `r length(unique(dat$id))` studies as of the
10\textsuperscript{th} of March 2021. Table \ref{tab:papers} contains the full
list of articles we analyse in this paper.

@@

Our study sample reflects the development of the literature. Although the "law
of $1/n$" was first formulated in 1981, the empirical assessment of the theory
only started a few years later, as dates of publishing range from `r min(dat$year)`
to `r max(dat$year)`. Most studies focus on the United States (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='USA')])`), but our sample also contains
papers on Australia (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='AUS')])`), Brazil (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='BRA')])`), Germany (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='DEU')])`), Indonesia (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='IDN')])`), Italy (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='ITA')])`), and
Switzerland (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='CHE')])`). Eight articles use cross-national data and analyse from 2 to 110
countries. Early studies used OLS and panel data methods to estimate the
results, while studies from 2012 onward have also applied causal inference
models such as instrumental variables and regression discontinuity designs to
test the relationship between house size and public spending.

@@

A key methodological issue we had to address concerns the potential violation
of an important assumption in a meta-analysis, that of effect size independence
[@cheung2014modeling; @cheung2019guide; @veroniki2016methods]. In our study
sample, authors frequently use the same datasets, and almost all papers fit more
than one regression with similar variables, what suggests that the assumption
above does not hold. We use two procedures to tackle this problem. First, we
created two sets of study coefficients to reduce the impact of
multicollinearity in our estimations. The first group includes only the most
rigorous models from each paper, that is, those estimated with the largest $n$,
most control variables, and fixed effects if the authors added them. If the
article employed a regression discontinuity design, we chose the coefficient
from the optimal bandwidth or from the intermediate one. This sample
encompasses `r dim(dat)[1]` estimates, as 
`r as.numeric(sum(table(table(dat$id))[-1]))` articles analysed two dependent or
independent variables of interest\footnote{The papers that used more than one
dependent or independent variable of interest are \citet{baqir1999districts,
bjedov2014impact, bradbury2001legislative, chen2007law, crowley2019law,
erler2007termlimits, gilligan2001fiscal, lee2016supermajority, lee2018court,
maldonado2013legislatures, primo2006stop, ricciuti2003trading,
ricciuti2004legislatures}.}. Our second sample, in contrast, contains all the
`r dim(fulldat)[1]` effect sizes reported in the `r length(unique(dat$id))`
papers. Here we focus on the results for our restricted sample as we consider
them more robust, but the findings are nearly identical when we use the extended
set.

Our second procedure consists of employing multilevel random effect models
[@cheung2014modeling; @matthes2019meta] in all of our estimations. We add two
extra levels to the regular meta-analysis, one including a unique publication
ID for each paper, and another indicating the data source used in the original
study. By adding these two levels, we account for within- and between-study
variation, thus removing these sources of effect size dependency and improving
the accuracy of the results. More information about the multilevel models can
be found in Section H of the Supplementary Material.

@@

extract_coefs <- function(mod, modnam) {
  dfr <- data.frame(
     nams2 = row.names(mod[["beta"]]),
     est = mod[["beta"]],
     se = mod[["se"]],
     pval = mod[["pval"]],
     model = modnam
  )
  aux <- data.frame(nams = c("Intercept",
             "Indepvar: logN",
             "Indepvar: N",
             "Year",
             "Elecsys: Non-Majoritarian",
             "Published",
             "Method: Panel",
             "Method: IV",
             "Method: RDD",
             "Legislature: Mixed",
             "Legislature: Unicameral"), 
             nams2 = c("intrcpt", 
                       "indepvar2logN", 
                       "indepvar2N", 
                       "year", 
                       "publishedYes", 
                       "elecsys2Non-Maj", 
                       "methodPANEL", 
                       "methodIV", 
                       "methodRDD",
                       "instdesignMixed",
                       "instdesignUnicameral"))
  dfr <- left_join(aux, dfr) %>%
    mutate_if(is.numeric, list(~round(., digits = 4)))
  dfr$nams2 <- NULL
  return(dfr)
}

gcf <- function(mod, pos) {
  x <- mod$est[pos]
  y <- mod$pval[pos]
  if(!is.na(x)) {
    return(paste0(round(x, digits = 3), strep(y)))
  } else {
    return('')
  }
}
gcf2 <- function(mod, pos) {
  x <- mod$se[pos]
  if(!is.na(x)) {
    return(paste0('(', round(x, digits = 3), ')'))
  } else {
    return('')
  }
}

# Expenditure per capita - Restricted
mod <-  rma(yi = coef,
              sei = SE,
              data = dat,
              method = "REML",
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
            subset = dat$depvar2=='ExpPC',
            slab = dat$authoryear,
            control=list(maxiter=1000))

@@

# Expenditure per capita - Extended
mod <- rma(yi = coef,
              sei = SE,
              data = fulldat,
              method = "REML",
              mods = ~ ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
            subset = fulldat$depvar2=='ExpPC',
            slab = fulldat$authoryear)

@@

# Log of Expenditure per capita - Restricted
mod <- rma(yi = coef,
              sei = SE,
              data = dat,
              method = "REML",
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
            subset = dat$depvar2=='logExpPC',
            slab = dat$authoryear)

@@

# Log of Expenditure per capita - Extended
mod <- rma(yi = coef,
              sei = SE,
              data = fulldat,
              method = "REML",
              mods = ~ ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
            subset = fulldat$depvar2=='logExpPC',
            slab = fulldat$authoryear)

@@

## Expenditure as percentage GDP -- Restricted
mod <-  rma.mv(yi = coef,
              V = VAR,
              data = dat,
              method = "REML",
              random = ~ 1 | id_level1/id_level2, 
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
              sparse = TRUE,
              tdist = TRUE,
              subset = dat$depvar2=='PCTGDP',
              slab = dat$authoryear)

@@

## Expenditure as percentage GDP -- Extended
mod <- rma(yi = coef,
              sei = SE,
              data = fulldat,
              method = "REML",
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
            subset = fulldat$depvar2=='PCTGDP',
            slab = fulldat$authoryear)

@@

\footnotetext{\textbf{Note:} {The restricted and extended samples include `r length(dat$id)` and `r length(fulldat$id)` study coefficients, respectively. We report the results from the permutation tests. Reference categories: Independent Variable $=$ Upper House Size; Published $=$ No; Method $=$ OLS, Inst. Design $=$ Bicameral. Significance codes: *** $p < 0.01$; ** $p < 0.05$; * $p < 0.10$. Blank cells mean that there is no sufficient data to estimate the paramater.}}

@@

The third and fourth columns use the natural logarithm of expenditure per
capita as the dependent variable. Among the coefficients in the restricted
sample, those in published studies are significantly smaller than those in
working papers\footnote{We find no evidence of publication bias in our models.
The funnel plots for all estimations are available Sections H and I of the
Supplementary Material.}. Another two moderators are negatively associated with
the outcome in our larger study pool. They both refer to estimation methods.
Studies that employ panel/fixed effects or regression discontinuity designs
(RDDs) have lower coefficients for log expenditure per capita if we take OLS as
the reference category.

Three variables are statistically significant in the third set of
meta-regressions, which include public expenditure as a percentage of GDP as
the dependent variable. Both in our restricted and in our extended samples,
recent studies have smaller coefficients than older papers, which stands in
contrast with the first model. Moreover, institutional design also affects
outcomes. Papers that use mixed samples of bicameral and unicameral systems
report significantly lower coefficients than those that analyse bicameral
legislatures exclusively. Non-majoritarian electoral systems are also
associated with lower levels of public spending, which may be justified since
the "law of $1/n$" was conceived for majoritarian voting. These latter results,
however, do not replicate in the first set of estimations.

The last two columns report meta-regressions that aggregate all selected
studies. In the restricted sample of coefficients, unicameralism has a positive
effect. This result holds for the extended sample as well. When we regress all
`r dim(fulldat)[1]` coefficients, the effects of estimation methods become
stronger once again. Panel/fixed-effects models and regression discontinuity
designs both significantly decrease the $1/n$ effect. Instrumental variable
models follow along these lines. Additionally, non-majoritarian voting
constituencies appear once again as settings where the "law of $1/n$" becomes
weaker.

>>>>

Aggregate results indicate that legislature size has no significant impact on
public spending. All of our main meta-analysis estimates show that the overall
effect is not statistically different from zero, thus confirming the
conflicting findings reported by the literature. However, when we look only at
articles that employ regression discontinuity designs (RDDs), we see that all
four papers included in our sample suggest that a higher number of legislators
leads to lower public spending [@debenedetto2018effect; @hohmann2017effect;
@lewis2019legislature; @petterssonlidbom2012size]. In this regard, it is
possible that methodological choices partially explain the divergent results
observed in the literature, as articles that use RDDs consistently point to the
same direction. One limitation of this finding is that these studies only test
the impact of lower house size and its natural logarithm on the natural logarithm of expenditure per
capita, thus it remains unclear whether the results hold with other variables.

The meta-regressions provide additional evidence that our study sample is
highly heterogeneous and that effect sizes differ substantially according to
study specifications. When using an extended sample of `r dim(fulldat)[1]`
coefficients, we find that unicameralism is associated with higher public
spending, as predicted by the "law of $1/n$". Moreover,
one of our meta-regressions indicates that larger upper chambers spend more in terms of
per capita expenditure than lower chambers, a result that also appears in the
binomial tests. Overall, non-majoritarian voting systems seem to decrease government spending, following the idea that the $1/n$ effect grows
weaker as the empirical cases distance from the original definition of the law.
Finally, meta-regression results confirm that regression discontinuity designs
reduce public spending estimates.

@@

In the second round, we did not require articles to cite @weingast1981political
and used a keyword-based query on Google Scholar to broaden the scope of the
first search. The search string contained terms strongly associated with the
literature on the "law of $1/n$" and it was as follows: `("upper chamber size" OR
"lower chamber size" OR "council size" OR "parliament size" OR "legislature
size" OR "number of legislators" OR "legislative size") AND ("spending" OR
"expenditure" OR "government size")`. We again restricted the search to
articles written in English which employed quantitative methods. This search
added two new results to our sample [@coate2011government;
@debenedetto2018effect], but neither of them included variables beyond the six
measures we had previously identified. In the third search round, we looked
into the personal webpages of every author whom we had already included in our
sample. The purpose of this manual search was to assess whether there was any
working paper or unpublished manuscript that we had missed in the two former
queries. We did not find any new article that satisfied the inclusion criteria
in this search. The full list of excluded records is available for online
consultation in the replication materials. Combined, the three searches
produced a dataset of `r length(unique(dat$id))` studies as of the
10\textsuperscript{th} of March 2021. Table \ref{tab:papers} contains the full
list of articles we analyse in this paper.

@@

Our study sample reflects the development of the literature. Although the "law
of $1/n$" was first formulated in 1981, the empirical assessment of the theory
only started a few years later, as dates of publishing range from `r min(dat$year)`
to `r max(dat$year)`. Most studies focus on the United States (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='USA')])`), but our sample also contains
papers on Australia (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='AUS')])`), Brazil (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='BRA')])`), Germany (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='DEU')])`), Indonesia (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='IDN')])`), Italy (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='ITA')])`), and
Switzerland (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='CHE')])`). Seven articles use cross-national data and analyse from 2 to 110
countries. Early studies used OLS and panel data methods to estimate the
results, while studies from 2005 onward have also applied causal inference
models such as instrumental variables and regression discontinuity designs to
test the relationship between house size and public spending.

@@

A key methodological issue we had to address concerns the potential violation
of an important assumption in a meta-analysis, that of effect size independence
[@cheung2014modeling; @cheung2019guide; @veroniki2016methods]. In our study
sample, authors frequently use the same datasets, and almost all papers fit more
than one regression with similar variables, what suggests that the assumption
above does not hold. We use two procedures to tackle this problem. First, we
created two sets of study coefficients to reduce the impact of
multicollinearity in our estimations. The first group includes only the most
rigorous models from each paper, that is, those estimated with the largest $n$,
most control variables, and fixed effects if the authors added them. If the
article employed a regression discontinuity design, we chose the coefficient
from the optimal bandwidth or from the intermediate one. This sample
encompasses `r dim(dat)[1]` estimates, as 
`r as.numeric(sum(table(table(dat$id))[-1]))` articles analysed two dependent or
independent variables of interest\footnote{The papers that used more than one
dependent or independent variable of interest are \citet{bjedov2014impact, bradbury2001legislative, chen2007law, crowley2019law,
erler2007termlimits, gilligan2001fiscal, lee2015supermajority, lee2016supermajority, lee2018court,
maldonado2013legislatures, primo2006stop, ricciuti2003trading,
ricciuti2004legislatures}.}. Our second sample, in contrast, contains all the
`r dim(fulldat)[1]` effect sizes reported in the `r length(unique(dat$id))`
papers. Here we focus on the results for our restricted sample as we consider
them more robust, but the findings are nearly identical when we use the extended
set.

Our second procedure consists of employing multilevel random effect models
[@cheung2014modeling; @matthes2019meta] in all of our estimations. We add two
extra levels to the regular meta-analysis, one including a unique publication
ID for each paper, and another indicating the data source used in the original
study. By adding these two levels, we account for within- and between-study
variation, thus removing these sources of effect size dependency and improving
the accuracy of the results. More information about the multilevel models can
be found in Section H.1 of the Supplementary Material.

@@

extract_coefs <- function(mod, modnam) {
  dfr <- data.frame(
     nams2 = row.names(mod[["beta"]]),
     est = mod[["beta"]],
     se = mod[["se"]],
     pval = mod[["pval"]],
     model = modnam
  )
  aux <- data.frame(nams = c("Intercept",
             "Indepvar: logN",
             "Indepvar: N",
             "Year",
             "Published",
             "Elecsys: Non-Majoritarian",
             "Method: Panel",
             "Method: IV",
             "Method: RDD",
             "Legislature: Mixed",
             "Legislature: Unicameral"), 
             nams2 = c("intrcpt", 
                       "indepvar2logN", 
                       "indepvar2N", 
                       "year", 
                       "publishedYes", 
                       "elecsys2Non-Maj", 
                       "methodPANEL", 
                       "methodIV", 
                       "methodRDD",
                       "instdesignMixed",
                       "instdesignUnicameral"))
  dfr <- left_join(aux, dfr) %>%
    mutate_if(is.numeric, list(~round(., digits = 4)))
  dfr$nams2 <- NULL
  return(dfr)
}

gcf <- function(mod, pos) {
  x <- mod$est[pos]
  y <- mod$pval[pos]
  if(!is.na(x)) {
    return(paste0(format(round(x, digits = 3), nsmall = 3), strep(y)))
  } else {
    return('')
  }
}
gcf2 <- function(mod, pos) {
  x <- mod$se[pos]
  if(!is.na(x)) {
    return(paste0('(', format(round(x, digits = 3), nsmall = 3), ')'))
  } else {
    return('')
  }
}

# Expenditure per capita - Restricted
mod <- rma.mv(yi = coef,
              V = VAR,
              data = dat,
              method = "REML",
              random = ~ 1 | id_level1/id_level2, 
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
              sparse = TRUE,
              tdist = TRUE,
              subset = dat$depvar2=='ExpPC',
              slab = dat$authoryear)

@@

# Expenditure per capita - Extended
mod <- rma.mv(yi = coef,
              V = VAR,
              data = fulldat,
              method = "REML",
              random = ~ 1 | id_level1/id_level2, 
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
              tdist = TRUE,
              sparse = TRUE,
              subset = fulldat$depvar2=='ExpPC',
              slab = fulldat$authoryear)

@@

# Log of Expenditure per capita - Restricted
mod <- rma.mv(yi = coef,
              V = VAR,
              data = dat,
              method = "REML",
              random = ~ 1 | id_level1/id_level2, 
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
              tdist = TRUE,
              sparse = TRUE,
              subset = dat$depvar2=='logExpPC',
              slab = dat$authoryear)

@@

# Log of Expenditure per capita - Extended
mod <- rma.mv(yi = coef,
              V = VAR,
              data = fulldat,
              method = "REML",
              random = ~ 1 | id_level1/id_level2, 
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
              tdist = TRUE,
              sparse = TRUE,
              subset = fulldat$depvar2=='logExpPC',
              slab = fulldat$authoryear)

@@

## Expenditure as percentage GDP -- Restricted
mod <- rma.mv(yi = coef,
              V = VAR,
              data = dat,
              method = "REML",
              random = ~ 1 | id_level1/id_level2, 
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
              tdist = TRUE,
              sparse = TRUE,
              subset = dat$depvar2=='PCTGDP',
              slab = dat$authoryear)

@@

## Expenditure as percentage GDP -- Extended
mod <- rma.mv(yi = coef,
              V = VAR,
              data = fulldat,
              method = "REML",
              random = ~ 1 | id_level1/id_level2, 
              mods = ~indepvar2+year+published+elecsys2+method+instdesign,
              test = "knha",
              tdist = TRUE,
              sparse = TRUE,
            subset = fulldat$depvar2=='PCTGDP',
            slab = fulldat$authoryear)

@@

\footnotetext{\textbf{Note:} {The restricted and extended samples include `r length(dat$id)` and `r length(fulldat$id)` study coefficients, respectively. We report the results from the permutation tests. Reference categories: Independent Variable $=$ Upper House Size; Published $=$ No; Method $=$ OLS, Inst. Design $=$ Bicameral. Significance codes: *** $p < 0.01$; ** $p < 0.05$; * $p < 0.10$. Blank cells mean that there is no sufficient data to estimate the parameter.}}

@@

The third and fourth columns use the natural logarithm of expenditure per
capita as the dependent variable. Among the coefficients in the restricted
sample, those in published studies tend to be smaller than those in
working papers\footnote{We find no evidence of publication bias in our models.
The funnel plots for all estimations are available Sections H and I of the
Supplementary Material.}. Another two moderators are negatively associated with
the outcome in our larger coefficient pool. They both refer to estimation methods.
Studies that employ panel/fixed effects or regression discontinuity designs
(RDDs) have lower coefficients for log expenditure per capita if we take OLS as
the reference category.

Three estimates are statistically significant in the third set of
meta-regressions, which include public expenditure as a percentage of GDP as
the dependent variable. Both in our restricted and in our extended samples,
recent studies have smaller coefficients than older papers, which stands in
contrast with the first model. Moreover, institutional design also affects
outcomes. Papers that use mixed samples of bicameral and unicameral systems
report significantly lower coefficients than those that analyse bicameral
legislatures exclusively. Non-majoritarian electoral systems, a category that encompasses proportional representation and mixed systems, may have a positive effect on the "law of $1/n$", but the estimate only borders on statistical significance.

The last two columns report meta-regressions that aggregate all selected
studies. In the restricted sample of coefficients, unicameralism has a positive
effect. This result holds for the extended sample as well. When we regress all
`r dim(fulldat)[1]` coefficients, the effects of estimation methods become
stronger once again. Panel/fixed-effects models and regression discontinuity
designs both significantly decrease the $1/n$ effect. Instrumental variable
models follow along these lines. Non-majoritarian electoral systems are also
significantly associated with lower levels of public spending, which may be justified since
the "law of $1/n$" was conceived for majoritarian voting. These latter results,
however, do not replicate in the other sets of estimations.
